Methods for Analyzing Models of\\ Rod-Shaped Bacteria

### 2025-12-04 paper/main.tex
add explanations for confidence intervals
minor style adjustments

\begin{figure}[H]
    \centering
    \begin{tikzonimage}[width=0.3\textwidth]
        {figures/crm_amir/progressions/step4-000010.png}
        \node at (0.02,17pt)[anchor=north west, white]{\textbf{A}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.3\textwidth]
        {figures/crm_amir/progressions/step4-000017.png}
        \node at (0.02,17pt)[anchor=north west, white]{\textbf{B}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.3\textwidth]
        {figures/crm_amir/progressions/step4-000034.png}
            \node at (0.01, 0.01+17pt)[anchor=north west, white]{\textbf{C}};
    \end{tikzonimage}\\
    \begin{tikzonimage}[width=0.33\textwidth]
        {figures/crm_amir/profiles-reduced/fit-comparison.pdf}%
        \node at (0.13, 0.88)[anchor=north west, black]{\textbf{B}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.33\textwidth]
        {figures/crm_amir/profiles/drag_force.pdf}%
        \node at (0.13, 0.88)[anchor=north west, black]{\textbf{D}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.33\textwidth]
        {figures/crm_amir/profiles/growth_rate.pdf}%
        \node at (0.13, 0.88)[anchor=north west, black]{\textbf{E}};
    \end{tikzonimage}\\
    \begin{tikzonimage}[width=0.33\textwidth]
        {figures/crm_amir/profiles/rod_rigidity.pdf}%
        \node at (0.13, 0.88)[anchor=north west, black]{\textbf{F}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.33\textwidth]
        {figures/crm_amir/profiles/spring_tension.pdf}%
        \node at (0.13, 0.88)[anchor=north west, black]{\textbf{G}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.33\textwidth]
        {figures/crm_amir/profiles/damping.pdf}%
        \node at (0.13, 0.88)[anchor=north west, black]{\textbf{H}};
    \end{tikzonimage}%
    \caption{
        (A-C) Manually segmented experimental snapshots of by Amir et al.~\cite{Amir2014_2}.
        The red line within the green mask shows the extracted position.
        (A-C) Snapshots of bending experiment by Amir et al.~\cite{Amir2014_2} with overlayed
        manual segmentation mask (bright green) and extracted position (red).
        The sequence shows the rod initially (A), under flow (B) and after relaxation (C).
        (D) Comparison of simulated and experimental rod in bent and relaxed state.
        (D-H) Profile-likelihoods for the estimated parameters.
        Brigth blue curves represent the full model, dark blue a reduced model without damping
        $\lambda=0$ and orange curves when fixing the auxiliary spring tension parameter
        $\gamma=\gamma_\text{opt}$.
        With these reductions, every parameter is identifiable.
    }
    \label{fig:amir-bending-simulation}
\end{figure}

@@

In this section we study a collection of growing, dividing and physically interacting
\ac{ecoli}.
We consider two scenarios to apply the previously developed metrics within a realistic context and
validate our approach.
Our dataset~\cite{Trude1982} consists of of a time-series of phase-contrast images (see
Figure~\ref{fig:application-2-data-overview} (A)).
We segmented them using omnipose~\cite{Cutler2022} (see Figure~\ref{fig:application-2-data-overview}
(B)) and extracted their trajectories with the previously described pipeline
(Section~\ref{sec:supplement-extraction-algorithm}).
The model is initialized with the extracted position of the first snapshot and parameters are
optimized to minimize the respective metrics.

@@

\begin{figure}[H]
    \centering
    \begin{tikzonimage}[width=0.49\textwidth]
        {figures/data/crm_fit/001044.png}%
        \node at (0.02, 0.98)[anchor=north west, white]{\textbf{A}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.49\textwidth]
        {figures/data/crm_fit/001044-markers.pdf}%
        \node at (0.02, 0.98)[anchor=north west, white]{\textbf{B}};
    \end{tikzonimage}\\
    \vspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.28\textwidth]
        {figures/crm_fit-data-progression.png}
        \node at (0.02, 0.98)[anchor=north west, white]{\textbf{C}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.35\textwidth]
        {figures/crm_estimate_params/IWF-Goettingen/rod-lengths-individual.pdf}%
        \node at (0, 1)[anchor=north west, black]{\textbf{D}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.35\textwidth]
        {figures/crm_estimate_params/IWF-Goettingen/growth-rates-bar-plot.pdf}%
        \node at (0, 1)[anchor=north west, black]{\textbf{E}};
    \end{tikzonimage}%
    \caption{
        \textbf{TODO adjust units in Plots (D-E)}
        (A) Snapshot of initial microscopic images.
        (B) Segmentation result of (A).
        (C) Initial (left), intermediate (middle) and final (right) microscopic snapshots.
        A complete list of snapshots is included in the supplementary material (see
        \nameref{sec:supplement-data-details}).
        (D) Analysis of growth for the chosen time interval.
        (E) Distribution of growth parameters shows cell-cell variability of the analyzed sample.
        The obtained results are further discussed in the supplementary material
        section~\ref{sec:supplement-reuse-estimated-parameters}.
    }
    \label{fig:application-2-data-overview}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_fit/profiles/damping.pdf}%
        \node at (0, 1.1)[anchor=north west, black]{\textbf{A}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_fit/profiles/exponent-m.pdf}%
        \node at (0, 1.1)[anchor=north west, black]{\textbf{B}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_fit/profiles/exponent-n.pdf}%
        \node at (0, 1.1)[anchor=north west, black]{\textbf{C}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_fit/profiles/strength.pdf}%
        \node at (0, 1.1)[anchor=north west, black]{\textbf{D}};
    \end{tikzonimage}\\
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_fit/profiles/radius.pdf}%
        \node at (0, 1.1)[anchor=north west, black]{\textbf{E}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_fit/profiles/growth-rate-0.pdf}%
        \node at (0, 1.1)[anchor=north west, black]{\textbf{F}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_fit/profiles/growth-rate-1.pdf}%
        \node at (0, 1.1)[anchor=north west, black]{\textbf{G}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_fit/profiles/growth-rate-2.pdf}%
        \node at (0, 1.1)[anchor=north west, black]{\textbf{H}};
    \end{tikzonimage}%
    \caption{
        (A-H) Likelihood profiles of optimized parameters.
        \textbf{TODO adapt this description to new results}
        Every plot indicates that a minimum was achieved, thus every parameter could be identified.
        We show only one of the 6 individually-chosen growth rate as an example.
        The remaining plots show similar behavior although with varying values and can be found in
        the supplementary material (Section~\ref{sec:supplement-parameter-profiles}).
    }
    \label{fig:parameter-estimates-single-step}
\end{figure}

>>>>

\begin{figure}[H]
    \centering
    \begin{tikzonimage}[width=0.3\textwidth]
        {figures/crm_amir/progressions/step4-000010.png}
        \node at (0.02,17pt)[anchor=north west, white]{\textbf{A}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.3\textwidth]
        {figures/crm_amir/progressions/step4-000017.png}
        \node at (0.02,17pt)[anchor=north west, white]{\textbf{B}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.3\textwidth]
        {figures/crm_amir/progressions/step4-000034.png}
            \node at (0.01, 0.01+17pt)[anchor=north west, white]{\textbf{C}};
    \end{tikzonimage}\\
    \begin{tikzonimage}[width=0.33\textwidth]
        {figures/crm_amir/profiles-reduced/fit-comparison.pdf}%
        \node at (0.13, 0.88)[anchor=north west, black]{\textbf{B}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.33\textwidth]
        {figures/crm_amir/profiles/drag_force.pdf}%
        \node at (0.13, 0.88)[anchor=north west, black]{\textbf{D}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.33\textwidth]
        {figures/crm_amir/profiles/growth_rate.pdf}%
        \node at (0.13, 0.88)[anchor=north west, black]{\textbf{E}};
    \end{tikzonimage}\\
    \begin{tikzonimage}[width=0.33\textwidth]
        {figures/crm_amir/profiles/rod_rigidity.pdf}%
        \node at (0.13, 0.88)[anchor=north west, black]{\textbf{F}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.33\textwidth]
        {figures/crm_amir/profiles/spring_tension.pdf}%
        \node at (0.13, 0.88)[anchor=north west, black]{\textbf{G}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.33\textwidth]
        {figures/crm_amir/profiles/damping.pdf}%
        \node at (0.13, 0.88)[anchor=north west, black]{\textbf{H}};
    \end{tikzonimage}%
    \caption{
        (A-C) Manually segmented experimental snapshots of by Amir et al.~\cite{Amir2014_2}.
        The red line within the green mask shows the extracted position.
        (A-C) Snapshots of bending experiment by Amir et al.~\cite{Amir2014_2} with overlayed
        manual segmentation mask (bright green) and extracted position (red).
        The sequence shows the rod initially (A), under flow (B) and after relaxation (C).
        (D) Comparison of simulated and experimental rod in bent and relaxed state.
        (D-H) Profile-likelihoods for the estimated parameters.
        Bands of blue and yellow indicate confidence intervals of $68\%,90\%$ and $95\%$.
        Brigth blue curves represent the full model, dark blue a reduced model without damping
        $\lambda=0$ and orange curves when fixing the auxiliary spring tension parameter
        $\gamma=\gamma_\text{opt}$.
        With these reductions, every parameter is identifiable.
    }
    \label{fig:amir-bending-simulation}
\end{figure}

@@

In this section we study a collection of growing, dividing and physically interacting
\ac{ecoli}.
We consider two scenarios to apply the previously developed metrics within a realistic context and
validate our approach.
Our dataset~\cite{Trude1982} consists of of a time-series of phase-contrast images (see
Figure~\ref{fig:application-2-data-overview} (A)).
We segmented them using the omnipose toolkit~\cite{Cutler2022} (see
Figure~\ref{fig:application-2-data-overview} (B)) and extracted their trajectories with the
previously described pipeline (Section~\ref{sec:supplement-extraction-algorithm}).
The model is initialized with the extracted position of the first snapshot and parameters are
optimized to minimize the respective metrics.

@@

\begin{figure}[H]
    \centering
    \begin{tikzonimage}[width=0.49\textwidth]
        {figures/data/crm_fit/001044.png}%
        \node at (0.02, 0.98)[anchor=north west, white]{\textbf{A}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.49\textwidth]
        {figures/data/crm_fit/001044-markers.pdf}%
        \node at (0.02, 0.98)[anchor=north west, white]{\textbf{B}};
    \end{tikzonimage}\\
    \vspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.28\textwidth]
        {figures/crm_fit-data-progression.png}
        \node at (0.02, 0.98)[anchor=north west, white]{\textbf{C}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.35\textwidth]
        {figures/crm_estimate_params/IWF-Goettingen/rod-lengths-individual.pdf}%
        \node at (0, 1)[anchor=north west, black]{\textbf{D}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.35\textwidth]
        {figures/crm_estimate_params/IWF-Goettingen/growth-rates-bar-plot.pdf}%
        \node at (0, 1)[anchor=north west, black]{\textbf{E}};
    \end{tikzonimage}%
    \caption{
        \textbf{TODO adjust units in Plots (D-E)}
        (A) Snapshot of initial microscopic images.
        (B) Segmentation result of (A).
        (C) Initial (left), intermediate (middle) and final (right) microscopic snapshots.
        A complete list of snapshots is included in the supplementary material (see
        \nameref{sec:supplement-data-details}).
        (D) Analysis of growth for the chosen time interval.
        (E) Distribution of growth parameters shows cell-cell variability of the analyzed sample.
        The obtained results are further discussed in the supplementary material
        (Section~\ref{sec:supplement-reuse-estimated-parameters}).
    }
    \label{fig:application-2-data-overview}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_fit/profiles/damping.pdf}%
        \node at (0, 1.1)[anchor=north west, black]{\textbf{A}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_fit/profiles/exponent-m.pdf}%
        \node at (0, 1.1)[anchor=north west, black]{\textbf{B}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_fit/profiles/exponent-n.pdf}%
        \node at (0, 1.1)[anchor=north west, black]{\textbf{C}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_fit/profiles/strength.pdf}%
        \node at (0, 1.1)[anchor=north west, black]{\textbf{D}};
    \end{tikzonimage}\\
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_fit/profiles/radius.pdf}%
        \node at (0, 1.1)[anchor=north west, black]{\textbf{E}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_fit/profiles/growth-rate-0.pdf}%
        \node at (0, 1.1)[anchor=north west, black]{\textbf{F}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_fit/profiles/growth-rate-1.pdf}%
        \node at (0, 1.1)[anchor=north west, black]{\textbf{G}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.25\textwidth]
        {figures/crm_fit/profiles/growth-rate-2.pdf}%
        \node at (0, 1.1)[anchor=north west, black]{\textbf{H}};
    \end{tikzonimage}%
    \caption{
        (A-H) Likelihood profiles of optimized parameters.
        Bands of blue and yellow represent confidence intervals of $68\%,90\%$ and $95\%$.
        \textbf{TODO adapt this description to new results}
        Every plot indicates that a minimum was achieved, thus every parameter could be identified.
        We show only one of the 6 individually-chosen growth rate as an example.
        The remaining plots show similar behavior although with varying values and can be found in
        the supplementary material (Section~\ref{sec:supplement-parameter-profiles}).
    }
    \label{fig:parameter-estimates-single-step}
\end{figure}
